---
title: "How do educated people kick the bucket?"
output: html_document
---

## Introduction

I'm interested in finding out what the differences in deaths are between education groups. I assume, that the more educated that one is, the less likely one will die by non-natural causes. 

First, the input data is loaded in.

```{r}
# Packages
library(readr)
library(dplyr)
library(ggplot2)

# Read input data
deathRecords = read_csv("input/DeathRecords.csv")
mannerOfDeath = read_csv("input/MannerOfDeath.csv")
education2003 = read_csv("input/Education2003Revision.csv")
```

Next, the dataset is subset and merged with the manner of death and education to determine the actual description of the data.

```{r}
filteredRecords = deathRecords[,c('Id', "MannerOfDeath", "Education2003Revision")]

# Merge together manner of death and educational info
filteredRecords = merge(filteredRecords, mannerOfDeath, by.x="MannerOfDeath", by.y="Code")
filteredRecords$MannerOfDeath = filteredRecords$Description
filteredRecords$Description = NULL
filteredRecords = merge(filteredRecords, education2003, by.x="Education2003Revision", by.y="Code")
filteredRecords$Education2003Revision = filteredRecords$Description
filteredRecords$Description = NULL
```

First, we take a quick look at the distribution of manner of deaths listed in the data.

```{r}
#Check the distribution of manner of death in the dataset
ggplot(filteredRecords, aes(x=MannerOfDeath, color=MannerOfDeath, fill=MannerOfDeath)) + 
  geom_bar(stat="count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("The Counts of the Manner of Deaths")
```

The majority of the deaths listed by a wide margin, appear to be natural deaths followed by not specified, accident, and suicide. 

Next, the data is aggregated based on education and manner of death to show the number of deaths for each manner of death in each education group.

```{r}
#Group by education and manner of death
grouped_data = aggregate(filteredRecords, 
                         by=list(filteredRecords$Education2003Revision, filteredRecords$MannerOfDeath),
                         FUN=length)

grouped_data$Id = NULL
grouped_data$MannerOfDeath = NULL
names(grouped_data) = c("Education", "MannerOfDeath", "Count")
grouped_data$Education = as.factor(grouped_data$Education)
grouped_data$MannerOfDeath = as.factor(grouped_data$MannerOfDeath)

```

```{r}
#Print bar graphs for the counts based on death type
for(i in levels(grouped_data$MannerOfDeath)){
  subset_level = subset(grouped_data, MannerOfDeath == i)
  titleCount = paste("The number of deaths per education for", i)
  print(titleCount)
  #Print counts
  print(ggplot(subset_level, aes(x=Education, y=Count, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(titleCount))
}
```


#Something about counts


An additional column is created to help determine the percentage of deaths per each education group.

```{r}
#Want to normalize to group to find percentage of count/total # in group
total8th = sum(grouped_data$Count[grouped_data$Education == "8th grade or less"])
total9th = sum(grouped_data$Count[grouped_data$Education == "9 - 12th grade, no diploma"])
totalAssoc = sum(grouped_data$Count[grouped_data$Education == "Associate degree"])
totalBac = sum(grouped_data$Count[grouped_data$Education == "Bachelor's degree"])
totalDoc = sum(grouped_data$Count[grouped_data$Education == "Doctorate or professional degree"])
totalHigh = sum(grouped_data$Count[grouped_data$Education == "high school graduate or GED completed"])
totalMast = sum(grouped_data$Count[grouped_data$Education == "Master's degree"])
totalColl = sum(grouped_data$Count[grouped_data$Education == "some college credit, but no degree"])
totalUnknown = sum(grouped_data$Count[grouped_data$Education == "Unknown"])

#Normalize data by dividing each value based on education group total numbers
grouped_data$PercentGroup = 0

for(i in 1:nrow(grouped_data)){
  if(grouped_data$Education[i] == "8th grade or less"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / total8th
  } else if(grouped_data$Education[i] == "9 - 12th grade, no diploma"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / total9th
  } else if(grouped_data$Education[i] == "Associate degree"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalAssoc
  } else if(grouped_data$Education[i] == "Bachelor's degree"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalBac
  } else if(grouped_data$Education[i] == "Doctorate or professional degree"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalDoc
  } else if(grouped_data$Education[i] == "high school graduate or GED completed"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalHigh
  } else if(grouped_data$Education[i] == "Master's degree"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalMast
  } else if(grouped_data$Education[i] == "some college credit, but no degree"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalColl
  } else if(grouped_data$Education[i] == "Unknown"){
    grouped_data$PercentGroup[i] = grouped_data$Count[i] / totalUnknown
  }
}
grouped_data$PercentGroup = grouped_data$PercentGroup * 100
```

```{r}
#Print bar graphs for the counts/percentage based on death type
for(i in levels(grouped_data$MannerOfDeath)){
  subset_level = subset(grouped_data, MannerOfDeath == i)
  titleCount = paste("The number of deaths per education of", i)
  titlePer = paste("The percentage of deaths per education of", i)
  #Print counts
  print(ggplot(subset_level, aes(x=Education, y=Count, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(titleCount))
  
  #Print percentage
  print(ggplot(subset_level, aes(x=Education, y=PercentGroup, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(titlePer))
}
```


```{r}
#Print bar graphs for the percentages based on death type
for(i in levels(grouped_data$MannerOfDeath)){
  subset_level = subset(grouped_data, MannerOfDeath == i)
  title = paste("The percentage of deaths per education for", i)
  print(ggplot(subset_level, aes(x=Education, y=PercentGroup, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(title))
}
```

```{r}
#Method to print both the count based and percentage based graphs
plotGraph = function(df, manner){
  subset_level = subset(df, MannerOfDeath == manner)
  titlePer = paste("The percentage of deaths per education for", manner)
  print(ggplot(subset_level, aes(x=Education, y=PercentGroup, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(titlePer))
  titleCount = paste("The number of deaths per education for", manner)
  print(ggplot(subset_level, aes(x=Education, y=Count, color=Education, fill=Education)) + 
    geom_bar(stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle(titleCount))
}
```