---
title: "How do educated people kick the bucket?"
output: html_document
---

## Introduction

This is a very light exploratory analysis primarily done to familiarize myself with the highcharter package.

First, the necessary libraries and datafiles are loaded.


## Preprocessing the data

```{r}
#Load necessary libraries
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library(highcharter)
library(dplyr)
```

```{r}
#Load datafile
vg <- read_csv("input/Video_Games_Sales_as_at_22_Dec_2016.csv")
```

Next, some very preliminary look into the data to see how the data is structured and looks like:

```{r}
head(vg)
```

```{r}
tail(vg)
```

As can be seen, the data is broken up based on the title of the games, platform, when it was released, genre, publishers, sales data, and ratings.

For now, we check to see if there are NA values in any of the columns.

```{r}
sapply(vg, function(x)(anyNA(x)))
```

As can be seen, not all for of the rows are completely filled in, including useful classifier data including the name or genre. To confirm what the data is comprised of, we will see the classes of all of the columns.

```{r}
str(vg)
```

As can be seen in the output, the year and the user score will have to be fixed to become numeric data columns.

```{r}
vg_platform <- vg %>% 
  count(Platform) %>% 
  glimpse()

hchart(vg_platform, "column", hcaes(x = Platform, y = n))
```
