---
title: "Did Shinya Yamanaka's Nobel Prize Award in 2012 change the research scores of his affiliated schools?"
output: html_document
---

##Introduction

In 2012, Shinya Yamanaka was awarded the Nobel Peace prize for his discovery of induced pluripotent stem cells(iPS cells). His research, revolutionized stem cell research due to the simplicity and strengths of iPS cells.

iPS cells are a type of stem cell coming from adult stem cells. They are reverted from adult cell back into a pluripotent stem cell stage in which it can turn into any cell type. These cells are revolutionary due to the fact that they are very similar to embryonic stem cells, bypassing the need for any controversy involved with collecting embryonic stem cells. As these cells are capable of giving rise to every other cell type, they are being used in numerous stem cell research including stem cell therapy and organogenesis.

The question now is, did Yamanaka's award change any of the research related scores in either the Times, Shanghai, or the CWUR rankings for either Kyoto Univeristy or University of California, San Francisco(UCSF)?

##Rankings
The rankings being looked at are from Times Higher Education World University Rankings, the Center for World University Rankings(CWUR), and the Shanghai Ranking. They are all similar in that they all look into similar factors including the research and number of publications from a school. They all have different weights to how they rank schools, with strong weights towards research.

The range of years for the data vary. Shanghai has the longest range, followed by Times, and finally CWUR. CWUR begins from 2012, as it was established that year. Times begins from 2011, likely due to their reorganization and re-evaluation of their metrics. 

###CWUR
CWUR is appears to be graded primarily based on the faculty and future prospects for the students attending school there. The relevant scores that will be looked at will be the following:

* Number of alumni who went on to win international awards/medals - 25%
* Number of faculty who have won international awards/medals - 25%
* Number of research papers in reputable journals - 5%
* Number of research papers in highly influential journals - 5%
* Number of highly cited research papers - 5%

###Shanghai
The relevant criteria that Shanghai rates it's scores that will be examined are the following:

* Alumni as Nobel Laureates/Fields medalist - 10%
* Staff as Nobel Laureates/Fields medalists - 20%
* Highly cited researchers in 21 broad subjects - 20%
* Papers published in Nature and Science(big science publications) - 20%

###Times
The relevant columns for Times and it's breakdown is the following:

* Research reputation/research income/papers and staff/public research income/total research income - 30%
* Citation impact based on citations - 32.5%
* International outlook(collaboration/international staff/student ratio) - 7.5%

##Results

The data comparing the relevant ratings are below:

```{r, echo=FALSE, warning=FALSE}
##
## Kaggle World Ranking Analysis
##

#Load libraries
library(ggplot2)

#Read data
input_dir = "input/"
cwurData = read.csv(paste0(input_dir, "cwurData.csv")) #CWUR rating
#ed_attain_sup = read.csv(paste0(input_dir, "educational_attainment_supplementary_data.csv")) #Population schooling
#ed_expen_sup = read.csv(paste0(input_dir, "education_expenditure_supplementary_data.csv")) #Funding for schools
school_country = read.csv(paste0(input_dir, "school_and_country_table.csv")) #Country/School table
shanghai = read.csv(paste0(input_dir, "shanghaiData.csv")) #Shanghai rating
times = read.csv(paste0(input_dir, "timesData.csv")) #Times rating

#Fix some of the names for the shanghai scores
school_country = rbind(school_country, data.frame(school_name = "University of California, San Francisco", country = "United States of America"))
shanghai$university_name[shanghai$university_name == "University of California-Berkeley"] = "University of California, Berkeley"

#Merge the shanghai data to get countries
##Lost datasets due to naming issues with school_country, as well as not in dataset
shanghai = merge(shanghai, school_country, by.x = "university_name", by.y = "school_name")

##
## Investigate the ratings of both Kyoto and UCSF in regards to Yamanaka's first paper on iPS cells and nobel prize
##

#Subset out data pertaining to UCSF and Kyoto Dai
kyoto_ucsf.shang = shanghai[shanghai$university_name == "University of California, San Francisco" | 
                        shanghai$university_name == "Kyoto University",]
kyoto_ucsf.cwur = cwurData[cwurData$institution == "University of California, San Francisco" | 
                             cwurData$institution == "Kyoto University",]
kyoto_ucsf.times = times[times$university_name == "University of California, San Francisco" | 
                             times$university_name == "Kyoto University",]

#Sort based on institute and year
kyoto_ucsf.shang = kyoto_ucsf.shang[order(kyoto_ucsf.shang$university_name,kyoto_ucsf.shang$year),]
kyoto_ucsf.cwur = kyoto_ucsf.cwur[order(kyoto_ucsf.cwur$institution, kyoto_ucsf.cwur$year),]
kyoto_ucsf.times = kyoto_ucsf.times[order(kyoto_ucsf.times$university_name, kyoto_ucsf.times$year),]
  
```

```{r, echo=FALSE, warning=FALSE}
#Look at the times rating for Kyoto University
ggplot(data=kyoto_ucsf.times, aes(x=year, group=university_name)) + 
    geom_line(aes(y=as.numeric(levels(kyoto_ucsf.times$international)[as.numeric(kyoto_ucsf.times$international)]), color="International Outlook")) + 
    geom_line(aes(y=research, color="Research"), size=1) + 
    geom_line(aes(y=citations, color="Citations"), size=1) + 
    labs(x="Year", y="Score", title="Times rating", size=1) +
    theme(
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size=14, face="bold"),
        legend.title = element_blank()
    )
```

The above graph shows the data from the Times Ranking for Kyoto University as University of California, San Francisco was not part of its data. As the Nobel Prize was awarded in 2012, there appears to be a downward trend until a slight resurgence the year of the prize. The score with citations rises from 2011-2012 and plateaus until 2015. The international ratings rises greatest from 2012-2013 and reaches a high point until 2015-2016.

```{r, echo=FALSE, warning=FALSE}
#Look at CWUR rating
ggplot(data=kyoto_ucsf.cwur, aes(x=year, col=institution)) + 
    geom_line(aes(y=publications, color="Publications"), size=1) + 
    geom_line(aes(y=influence, color="Influence"), size=1) + 
    geom_line(aes(y=citations, color="Citations"), size=1) + 
    geom_line(aes(y=broad_impact, color="Broad impact"), size=1) +
    labs(x="Year", y="Score", title="CWUR rating") +
    theme(
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size=14, face="bold")
    ) + 
    facet_grid(institution~.,scales="free_x")
```

The above graph shows the scores taken from CWUR between 2012 and 2015 for both Kyoto University and UCSF. The scores vary depending on the universities. Part of broad impact is missing due to issues with the data. In both schools, citations appear to raise from 2012-2013. In UCSF, it begins to taper off after 2013 and begins to decline while Kyoto sharply raises again after 2014. The scores for publications both rise a bit after 2012 with higher scores seen in Kyoto Unversity. In 2012-2013, an increase in the scores for influence can be seen in Kyoto University. The metrics given by CWUR deems UCSF to not nearly be as influential as Kyoto University.

```{r, echo=FALSE, warning=FALSE}
#Look at Shanghai data
ggplot(data=kyoto_ucsf.shang, aes(x=year, col=university_name)) + 
    geom_line(aes(y=hici, group=university_name, color="Highly Cited"), size=1) + 
    geom_line(aes(y=ns, group=university_name, color="Nature/Science"), size=1) + 
    geom_line(aes(y=pub, group=university_name, color="Publication"), size=1) + 
    geom_line(aes(y=award, group=university_name, color="Num. of Award"), size=1) +
    labs(x="Year", y="Score", title="Shanghai rating") +
    theme(
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        plot.title = element_text(size=14, face="bold")
    ) + 
    facet_grid(university_name~.,scales="free_x")
```

Shanghai is markedly different as it contains data all the way back in 2005. The original paper published by Yamanaka was released in 2006. In both universities, the publication scores are trending downwards in both the release year of the publication as well as after the Nobel Prize. Kyoto University appears to have an upward trend in highly cited scores the year of the release of the paper with a dip about after the Nobel Prize. UCSF follows a gradual downward trend followed by a sharp decrease in 2013-2014. There appears to be a small jump in the award score following 2012. The number of papers in Nature or Science does not appear different due to the paper.


Yamanaka's reward appears to have affected the scores related in research in both CWUR and Times. The data from Shanghai's rating appeared to show no major changes to the scores after his Nobel Prize. The data overall from Shanghai appear to be consistent throughout the years. It should be noted that a single person is unlikely to be able to drastically change the scores for any of the college rankings and should be considered correlation.