---
title: "Kaggle - TalkingData Mobile User Demographics"
output: html_document
---

```{r}
# Packages
library(readr)
library(dplyr)

# Read input data
app_events = read_csv("input/app_events.csv")
app_labels = read_csv("input/app_labels.csv")
events = read_csv("input/events.csv")
labels = read_csv("input/label_categories.csv")
phone_brand_model = read_csv("input/phone_brand_device_model.csv")
test = read_csv("input/gender_age_test.csv")
train = read_csv("input/gender_age_train.csv")
```


```{r}
head(app_events)
```

Linked to events. Shows if app is installed and active.

```{r}
head(app_labels)
```

Shows the labels id for each app_id.

```{r}
head(labels)
```

Shows the label to cateogry for each app.

```{r}
head(events)
```

Data showing device id, to app with timestamp, lat long.

```{r}
head(phone_brand_model)
```

Need to consider looking into changing encoding for chinese characters. Other than that, device id to phone brand and device model.

```{r}
head(test)
head(train)
```

Goal is to predict the probability of each device id is for each gender/age group

