---
title: "What are the nationalities of the authors of the papers published for NIPS2015?"
output: html_document
---

## Introduction

Neural Information Processing Systems (NIPS) is one of the top machine learning conferences in the world. My question is, in general, which nationalities are represented in the event as authors of research papers? This is a generalization and done just for fun as there are many things that can easily affect the nationality or obscure whatever it could be.

Anyways, libraries and the data are first initialized and loaded.

```{r, message=FALSE}
# Packages
library(readr)
library(dplyr)
library(ggplot2)

# Read input data
authors = read_csv("input/Authors.csv")
paperAuthors = read_csv("input/PaperAuthors.csv")

#Function to find last name for each row
findLastName = function(fullname){
  return(fullname[length(fullname)])
}
```

## Authors of papers

The authors dataset contains the authors that have wrote papers. First, we take a quick look:

```{r}
#Check last names
authorNames = strsplit(authors$Name, " ")

lastNames = lapply(authorNames, findLastName)
lastNames = as.character(lastNames)

#Count number of each last name
lastNamesCount = as.data.frame(table(lastNames))
lastNamesCount = lastNamesCount[order(lastNamesCount$Freq),]
```

```{r}
#Load barplot of last names given in authorNames
ggplot(lastNamesCount, aes(factor(Freq), fill=factor(Freq))) + geom_bar() + 
  scale_fill_brewer(palette="Spectral") + 
  ggtitle("Number of unique surnames vs Number of Authors") + 
  xlab("Number of unique surnames") + 
  ylab("Number of authors")
```

As can be seen by the distribution, the distribution of unique surnames in the conference is right skewed with most authors' surnames being unique in the conference.

Now to see the top 10 author surnames.

```{r}
tail(lastNamesCount, 10)
```

The top surnames who had published a paper in this conference are Wang, Li, and Chen.

The question at this point is, what could the nationalities of the top 10 authors be?

The origins was manually curated to guess the nationality of the top 10 authors who had published papers for the event. The origins were found using wikipedia and deriving it from origin tabs for surnames. 

```{r, echo=FALSE}
lastNameOrigins = data.frame(
    lastName = tail(lastNamesCount$lastNames,10),
    origin = c("Chinese", "Chinese", "Chinese", "Chinese", "Chinese/English/Korean", "Chinese", "Chinese", "Chinese", "Chinese", "Chinese")
)
```

As can be seen from the list, the most popular surnames that have published papers for NIPS2015 are Chinese. Lee is also a common surname in Korea and in England so there can be some ambiguity there.

Now the question is, if there is any difference based on the numbers of papers that they have published.

## Authors and their published works

Now, we take a look into the authors and the number of papers they have published.

```{r}
#Find who wrote papers
publishers = merge(authors, paperAuthors, by.x = "Id", by.y="AuthorId")

#Check last names
authorNames = strsplit(publishers$Name, " ")
lastNamesPub = lapply(authorNames, findLastName)
lastNamesPub = as.character(lastNamesPub)

#Count number of each last name
lastNamesPubCount = as.data.frame(table(lastNamesPub))
lastNamesPubCount = lastNamesPubCount[order(lastNamesPubCount$Freq),]
```

```{r}
ggplot(lastNamesPubCount, aes(factor(Freq), fill=factor(Freq))) + geom_bar() + 
  scale_fill_brewer(palette="Spectral") + 
  ggtitle("Number of published papers vs Number of Authors") + 
  xlab("Number of papers published under a surname") + 
  ylab("Number of authors")
```

In this plot, the distribution for the number of papers published under a surname is larger. It is still right skewed but the number of surnames published a single paper drops while the number of of multiple surname rises. 

What about the change in nationality then?

```{r}
tail(lastNamesPubCount,10)
```

This list shows the number of a given surname that had published papers for this event. As can be seen in this list, the top publishers are Wang, Li, and Zhang. The lowest frequency becomes 7. The biggest jumps appear to be Ravikumar, which had published 7 papers and Song which published an average of 2.33 papers for 3 individuals.

```{r, echo=FALSE}
lastNamePubOrigins = data.frame(
    lastName = tail(lastNamesPubCount$lastNames,10),
    origin = c("Chinese", "Indian", "Chinese/Korean/African", "Chinese", "Chinese/English/Korean", "Chinese", "Chinese", "Chinese", "Chinese", "Chinese")
)
```

##Conclusion

It appears that Asians publish the most for this event. The surnames of the top publishers are very common surnames in China so it should not be a suprise that those surnames top the list. It would be interesting to see a breakdown based on everyone who participated but unforunately, there are no easy databases to query and would require a fair amount of work to manually annotate each surname if they can be found as well as the large amount of ambiguity that surname origins can bring. 