---
title: "What are the nationalities of the writers"
output: html_document
---

## Introduction

Neural Information Processing Systems (NIPS) is one of the top machine learning conferences in the world. I was wondering, in general, which nationalities are represented in the event as authors of research papers? This is a generalization and done just for fun.

```{r}
# Packages
library(readr)
library(dplyr)

# Read input data
authors = read_csv("input/Authors.csv")
paperAuthors = read_csv("input/PaperAuthors.csv")

#Function to find last name for each row
findLastName = function(fullname){
  return(fullname[length(fullname)])
}
```

##Authors of papers

The authors dataset contains the authors that have wrote papers. First, we take a quick look:

```{r}
#Check last names
authorNames = strsplit(authors$Name, " ")

lastNames = lapply(authorNames, findLastName)
lastNames = as.character(lastNames)

#Count number of each last name
lastNamesCount = as.data.frame(table(lastNames))
lastNamesCount = lastNamesCount[order(lastNamesCount$Freq),]

lastNamesCount
```


The origins was manually curated to guess the nationality of each author. The origins was created using a mixture of hits from wikipedia as well as through searches such as through ancestry.com. 

```{r, echo=FALSE}
origins = data.frame(
  lastName = lastNames,
  origin = c("Chinese", "Spanish", "German/Chinese", "Chinese", "German", "Ireland")
)
```

As can be seen from the list, the most popular surnames that have published papers are Wang, Li, and Chen. 

##Authors and their published works

Now, we take a look into the authors and the number of papers they have published.

```{r}
#Find who wrote papers
publishers = merge(authors, paperAuthors, by.x = "Id", by.y="AuthorId")

#Check last names
authorNames = strsplit(publishers$Name, " ")
lastNamesPub = lapply(authorNames, findLastName)
lastNamesPub = as.character(lastNamesPub)

#Count number of each last name
lastNamesPubCount = as.data.frame(table(lastNamesPub))
lastNamesPubCount = lastNamesPubCount[order(lastNamesPubCount$Freq),]

lastNamesPubCount
```

This list shows the number of a given surname that had published papers for this event. As can be seen in this list, the top publishers are Wang, Li, and Zhang.