---
title: "Who are the bootcampers"
output: html_document
---

I'm interested in seeing who are the types of people who attend bootcamps, and the impact of programming bootcamps. I have always felt that bootcamps, with the exceptions of a few of them, feel like a scam due to the lack of transparency and purported, high employment rates afterwards. Hopefully, this self-reported data can help answer some of those questions

First, the data is loaded.

```{r}
#Load libraries
library(ggplot2)
library(knitr)
library(plyr)

#Read the data
survey = read.csv("input/2016-FCC-New-Coders-Survey-Data.csv")
```

```{r}
set.seed(1)
sample(names(survey),20)
```

Judging from the names of some of the variables of the data, there are numerous types of data available such as the resources someone uses to learn, what they did in school, their current work, what they want to do, if they have attended bootcamps, etc.

Now let's take a look at those who had attended bootcamps.

```{r}
bootcampers = survey[which(survey$AttendedBootcamp == 1),]
```

It seems a little more than `r round((nrow(bootcampers)/nrow(survey)) * 100,2)`% of the surveyed demographics have attended bootcamps.

Let's take a look at some of the demographic data.

###Age

```{r}
hist(bootcampers$Age, main="Histogram of the Age Distribution of People who attended Bootcamps",
     xlab = "Age")
```

The distribution of the ages are right skewed. It is mostly younger people who are trying to attend. The average age of a person who went to a bootcamp is `r round(mean(bootcampers$Age, na.rm = T)) `.

###Debt

`r round(nrow(bootcampers[bootcampers$HasDebt == 1,])/nrow(bootcampers) * 100, 2) `% of the students has some form of debt.

`r round(nrow(bootcampers[bootcampers$HasStudentDebt == 1,])/nrow(bootcampers) * 100, 2) `% of the students have student debt.


```{r}
hist(bootcampers$StudentDebtOwe, main="Histogram of the Amount of Student Debt Owed",
     xlab="Student Debt Owed")
```

The amount of debt that people have is right skewed. Most of the surveyed people have student debts up to 50k with it sharply dropping after. It is possible that the very end value is a outlier as it is a bit over 3x more than the next value but it is a possible value if say, the person went into med school. To check if it is an outlier though, we can see if it is above 1.5 * IQR(interquartile range) of the third quartile or 1.5 * IQR below the 1st quartile.

```{r}
quantile(bootcampers$StudentDebtOwe, na.rm=TRUE)
10000 - 1.5 * IQR(bootcampers$StudentDebtOwe, na.rm=TRUE)
40000 + 1.5 * IQR(bootcampers$StudentDebtOwe, na.rm=TRUE)
```

The outliers are any value below -35000 and above 85000. As the lowest value is 2, any value about 85000 is considered an outlier.

```{r}
hist(bootcampers$StudentDebtOwe[bootcampers$StudentDebtOwe < 85000], main="Histogram of the Amount of Student Debt Owed", xlab="Student Debt Owed")
```

In redrawing the graph without the outliers, we can see the end of the data. We can see that the frequency has a decreasing trend until the small spike past the 70k mark.

###Education

```{r}
bootcampDegrees = as.data.frame(table(bootcampers$SchoolDegree))
names(bootcampDegrees) = c("Degree", "Freq")
bootcampDegrees = bootcampDegrees[order(bootcampDegrees$Freq),]
kable(bootcampDegrees)
```

The majority of people who attended bootcamps already had bachelor's degrees. Following them by a quarter are those with some college credit and then those with master's degrees.

What are the majors of the people then?

```{r}
bootcampMajors = as.data.frame(table(bootcampers$SchoolMajor))
bootcampMajors = bootcampMajors[bootcampMajors$Freq != 0,]
bootcampMajors = tail(bootcampMajors[order(bootcampMajors$Freq),], 10)
names(bootcampMajors) = c("Major", "Freq")
kable(bootcampMajors)
```

The majority of degree holders held English degrees followed by Computer Science and Psychology.

###Job Prospects

```{r}
fullJobAfter = as.data.frame(table(bootcampers$BootcampFullJobAfter))
names(fullJobAfter) = c("FullJobAfter", "Freq")
fullJobAfter$FullJobAfter = c("No", "Yes")
kable(fullJobAfter)
```
It seems like that more than half of the people who attended have gotten full time jobs afterwards. 

```{r}
hist(bootcampers$BootcampMonthsAgo[which(bootcampers$BootcampFullJobAfter == 1)], 
     main = "Histogram of the Number of Months after attending a Bootcamp before Full Time Job",
     xlab  = "Number of months ago after Attending Bootcamp")
```

According to the histogram, the majority of the people who attended, got jobs within a year of attending. There do seem to be outliers that may be due to false reporting as it is unlikely there were tech bootcamps back in 2003, especially a few years after the dot-com bubble crashed.

```{r}
jobAfterDegrees = as.data.frame(table(bootcampers$SchoolDegree[bootcampers$BootcampFullJobAfter == 1]))
names(jobAfterDegrees) = c("Degree", "Freq")
jobAfterDegrees = jobAfterDegrees[order(jobAfterDegrees$Freq),]
kable(jobAfterDegrees)
```

The bulk of the people who attended bootcamps and had jobs after bootcamp, already had bachelor's degrees. The next biggest group was those with master's degrees and finally those with some college credit.

```{r}
majors_hired = as.data.frame(table(bootcampers$SchoolMajor[bootcampers$BootcampFullJobAfter == 1]))
majors_hired = majors_hired[majors_hired$Freq != 0,]
majors_hired = tail(majors_hired[order(majors_hired$Freq),], 10)
names(majors_hired) = c("Major", "Freq")
kable(majors_hired)
```

The majority of those who were hired came from English majors, followed by political science, psychology, and economics majors. There do appear to be those who had degrees in IT and Computer Science, and other more directly computationally intensive majors listed here.

```{r}
table(bootcampers$SchoolDegree[bootcampers$BootcampFullJobAfter == 0])
```

The degrees held by those that did not manage to get a full time job after are similar to those who did. Bachelor's degree holders form the bulk of the numbers followed by master's degree holders. 

```{r}
majors_nothired = as.data.frame(table(bootcampers$SchoolMajor[bootcampers$BootcampFullJobAfter == 0]))
majors_nothired = majors_nothired[majors_nothired$Freq != 0,]
majors_nothired = tail(majors_nothired[order(majors_nothired$Freq),], 10)
names(majors_nothired) = c("Major", "Freq")
kable(majors_nothired)
```

The majors that was not hired the most was also English. Interestingly, it is followed by degree holders of Computer Science and Business Administration. 


```{r}
resources = as.data.frame(table(survey$ResourceOther))
names(resources) = c("Resource", "Freq")
resources = tail(resources[order(resources$Freq),], 10)
#Clean up duplicates from top 10
resources$Resource[resources$Resource %in% c("Treehouse ", "treehouse", "Team Treehouse")] = "Treehouse"
resources$Resource[resources$Resource %in% c("Lynda.com")] = "Lynda"

resources = ddply(resources, .(Resource), colwise(sum))
resources = resources[order(resources$Freq),]
kable(resources)
```

There were a variety of data resources that were used. A lot of the entries were not filed out or needed to be cleaned up. The top 10 resources which were cleaned up after the fact, were Youtube, w3, Stack Overflow, Books, Lynda, and Treehouse. The most popular, Treehouse, offers a huge variety of courses from web design to game development. Lynda is an older company which offers a huge variety and was recently bought by linkedin.

```{r}
bootcamps = as.data.frame(table(bootcampers$BootcampName))
names(bootcamps) = c("Bootcamp" , "Freq")
bootcamps = tail(bootcamps[order(bootcamps$Freq),], 10)
```

As there are a lot of different bootcamps, only the top 10 was examined as there were many with a single person attending. There are also some dummy data stemming from Free Code Camp not being an actual bootcamp.

General Assembly had the highest attendance rate. General Assembly is one of the oldest bootcamps as well as has numerous courses. Flatiron School is the second most attended and offers in person programs in New York as well as online. Dev Bootcamp is another one of the older bootcamps, offering courses in numerous places as well as having online courses now. 

The top three offers online courses as well as numerous other in person courses in many locations. The courses that they offer in common are web development related.

