---
title: "Exploratory"
output: html_document
---

```{r}
#Load libraries
library(dplyr)
library(ggplot2)
library(googleVis)
library(lubridate)
library(data.table)

#Read the data
zika = read.csv("input/cdc_zika.csv")
```

```{r}
#Clean up the 7 na rows
zika = zika[!(zika$location == ""),]
zika$location = droplevels(zika$location)

#Remove time_period/time_period_type variable
zika$time_period = NULL
zika$time_period_type = NULL
```

```{r}
#Preprocessing
#Clean up location to get country/area

#Splits location based on - for country
split_country = function(x){
  splitted = strsplit(as.character(x),"-")
  return((splitted[[1]][1]))
}

#Splits location based on - for province?/state
split_province = function(x){
  splitted = strsplit(as.character(x),"-")
  if(length(splitted[[1]]) > 1){
    return(paste(splitted[[1]][2:length(splitted[[1]])], collapse = "-"))
  } else {
    return(splitted[[1]][1])
  }
}

countries = tapply(levels(zika$location), levels(zika$location), split_country)
provinces = tapply(levels(zika$location), levels(zika$location), split_province) #Still seemingly has state-county

#Join table for locations
area_table = data.frame(Primary = dimnames(countries)[[1]], country = as.character(countries), provinces = as.character(provinces))

zika = inner_join(zika, area_table, by=c("location" =  "Primary"))
```


```{r}
table(zika$location)
```
Seems to be South America and parts of the US. Seems to have 7 missing locations.

```{r}
table(zika$report_date)
```

Looks like it's relatively recent data, being from late 2015 to very early July 2016. 

```{r}
table(zika$location_type)
```
Location type wise, it appears zika is mostly in minicipalities. There are 7 blank entries to take note of.

```{r}
table(zika$time_period)
table(zika$time_period_type)
```
Seems like the time period variables have no data.

```{r}
table(zika$unit)
```
There are 7 blank values again. The units appear to be split up into cases and municipalities.

```{r}
table(zika$data_field)
```

Seems to show results of each case municipality. The types seem to range from possible zika, zika suspicion, birth defects, and confirmations.

```{r}
muni_data = as.data.frame(table(zika$data_field[zika$unit == "municipalities"]))
muni_data[order(muni_data$Freq),]
as.data.frame(table(zika$location_type[zika$unit == "municipalities"]))
as.data.frame(table(zika$country[zika$unit == "municipalities"]))
```

For the reported incidents that are in municipalities, each of them suspect microcephaly or have cases of it. In addition, they are all reported in Brazil in states.

```{r}
table(zika$country)
table(zika$province)
```

There appears to be Spanish direction words also mixed into location such as Norte, Centro, and Nordeste. There oddly appears to be 7 incidents in each of those directions.

```{r}
head(zika[zika$country == "United_States",])
```

The cases in the US started near the end of Feburary in 2016.

```{r}
#Plot the country/incident count
countries = zika[!(zika$country %in% c("Centro", 'Nordeste', 'Norte', 'Sudeste', 'Sul')),]
countries$country = gsub("_", " ", countries$country)
ggplot(countries, aes(x=countries$country, fill=countries$country)) + 
  geom_bar() + 
  labs(x= "Country", y = "Count", title = "Number of incidents per country") +
  scale_fill_discrete(name = "Country") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        text = element_text(size=15))
```

Zika cases have been primarily reported in South America, in particular, Colombia. The most number of incidents reported comes from the United States Virgin Islands, followed by El Salvador.

```{r}
#Process in date info
zika$Year = NA
zika$Month = NA
zika$Day = NA

zika$Year = year(as.Date(zika$report_date))
zika$Month = month(as.Date(zika$report_date))
zika$Day = day(as.Date(zika$report_date))
```

```{r}
#Plot data with googleVis
#Suffers from max 400 hits via googlevis
plot_map = function(startDay, startMonth, startYear, endDay, endMonth, endYear){
  if(startYear <= endYear){
    if(startMonth <= endMonth){
      timePeriod = subset(zika, Year <= endYear & Year >= startYear)
      timePeriod = subset(timePeriod, Month <= endMonth & Month >= startMonth)
      timePeriod = subset(timePeriod, Day <= endDay & Day >= startDay)
      plot(gvisMap(timePeriod, "location", "report_date"))
    }
  }
}
```