---
title: "Exploratory"
output: html_document
---

```{r}
#Read the data
zika = read.csv("input/cdc_zika.csv")
```

```{r}
table(zika$location)
```
Seems to be South America and parts of the US. Seems to have 7 missing locations.

```{r}
table(zika$report_date)
```

Looks like it's relatively recent data, being from late 2015 to very early July 2016. 

```{r}
table(zika$location_type)
```
Location type wise, it appears zika is mostly in minicipalities. There are 7 blank entries to take note of.

```{r}
table(zika$time_period)
table(zika$time_period_type)
```
Seems like the time period variables have no data.

```{r}
table(zika$unit)
```
There are 7 blank values again. The units appear to be split up into cases and municipalities.

```{r}
table(zika$data_field)
```

Seems to show results of each case municipality. The types seem to range from possible zika, zika suspicion, birth defects, and confirmations.

```{r}
#Clean up the 7 na rows
zika = zika[!(zika$location == ""),]
levels(zika$location) = droplevels(zika$location)
```

```{r}
#Clean up location to get country/area
for(i in 1:nrow(zika)){
  splitted = strsplit(as.character(zika$location[i]),"-")[[1]]
  zika$Country[i] = splitted[1]
  if(length(splitted) > 1){
    zika$Area[i] = paste(splitted[2:length(splitted)], collapse = "-")
  } else {
    zika$Area[i] = splitted[1]
  }
}

#Splits location based on - for country
split_country = function(x){
  splitted = strsplit(as.character(x),"-")
  return((splitted[[1]][1]))
}

#Splits location based on - for province?/state
split_province = function(x){
  splitted = strsplit(as.character(x),"-")
  if(length(splitted[[1]]) > 1){
    return(paste(splitted[[1]][2:length(splitted[[1]])], collapse = "-"))
  } else {
    return(splitted[[1]][1])
  }
}

countries = tapply(levels(zika$location), levels(zika$location), split_country)
provinces = tapply(levels(zika$location), levels(zika$location), split_province)
```

```{r}
#Load libraries
library(ggplot2)
library(maps)

#Load state data
all_states = map_data("state")

#Plot all states with ggplot
p = ggplot() + geom_polygon( data=all_states, aes(x=long, y=lat, group = group),colour="white", fill="grey10" )

```