---
title: "New User Bookings Exploratory Analysis"
output: html_document
---

This is just some quick exploratory analysis done on the data provided. Anyways, the data is first loaded.

```{r}
#Packages
library(ggplot2)
library(grid)
library(gridExtra)

#Data
dir = 'input/'
train = read.csv(paste0(dir,"train_users.csv"))
test = read.csv(paste0(dir,"test_users.csv"))
countries = read.csv(paste0(dir,"countries.csv"))
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
sessions  = read.csv(paste0(dir,"sessions.csv"))
```

The training and testing data is merged together to look at the data that will be used for prediction

```{r}
#needs to do merge all
train = merge(train, sessions, by.x="id", by.y="user_id")
test = merge(test, sessions, by.x="id", by.y="user_id")
```

#Training/Testing data

Create a quick function to show the head, tail, and number of na for each variable called

```{r}
sapply(train, function(y) sum(length(which(is.na(y)))))
```

The only column with NA values is age in train.

```{r}
sapply(test, function(y) sum(length(which(is.na(y)))))
```

The same can be seen in the test data set.

```{r}
table(train$age)
```

```{r}
table(test$age)
```

As can be seen in both the test and train set, the age variable will need some cleaning up. It is extremely unlikely that anyone 5 and under actually signed up and used the service. Because technically the rules state that you must be 18 to use the service, ages under 18 should be cleaned up. In addition, it appears that the ages variable has birth year data so those should be cleaned up accordingly. Although it would be hard to verify ages 100+, the oldest human recorded was around 117 so the outlier should be changed as well.

```{r}
test$age[test$age > 150 & !is.na(test$age)] = 2015 - test$age[test$age > 150 & !is.na(test$age)]
train$age[train$age > 150 & !is.na(train$age)] = 2015 - train$age[train$age > 150 & !is.na(train$age)]
test$age[test$age > 115 & !is.na(test$age)] = 115
train$age[train$age > 115 & !is.na(train$age)] = 115
test$age[test$age < 18 & !is.na(test$age)] = 18
train$age[train$age < 18 & !is.na(train$age)] = 18

train_ages = ggplot(train,aes(age)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Distribution of ages in train")
test_ages = ggplot(test,aes(age)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Distribution of ages in test")

grid.arrange(train_ages, test_ages, ncol = 2, main = "Main title")

```

#Sessions
```{r}
#See which values have na
sapply(sessions, function(y) sum(length(which(is.na(y)))))
```

The only columns with NA values is secs_elapsed.

```{r}
##Analyze the na sec
na_secs = sessions[which(is.na(sessions$secs_elapsed)),]
```

```{r}
##go take the top 10 actions
ggplot(sessions,aes(action)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```



```{r}
##look into the blank actiontype
#See the counts of action types
ggplot(sessions,aes(action_type)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

The most popular action type is views, followed by data and click.

```{r}
##take the top 10
ggplot(sessions,aes(action_detail)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

```{r}
#Show the counts of device types
ggplot(sessions,aes(device_type)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Counts of device types")
```



```{r}
ggplot(sessions,aes(secs_elapsed)) + geom_dotplot() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Counts of device types")
```


#Age data

The age data shows age bucket groups of users going to certain country destinations in 2015. 

```{r}
#switch 100+ age to 99+
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
levels(age$age_bucket) = c(levels(age$age_bucket), "99+")
age$age_bucket[age$age_bucket == "100+"] = "99+"

#plot destination vs population with gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Total population in thousands going to destinations separated by gender")
```

This shows that the gender distribution towards countries are roughly the same. The most population destination is within the US with the second most being Germany.

```{r}
#plot age vs population
ggplot(age,aes(x=age_bucket, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in the graph, there is an increase in travelers using the service in their 20s and begins to dip in the ages of 30-50. There is an upsurge in travel again in the travelers that are in their 50s and and begins to steadily decrease with age. There appears to be more males using the service that are in their mid-30s and below however begins to even out and become eclipsed by their female counterpart afterwards.

```{r}
#plot destinatoin vs population with alpha of age and gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender, alpha=age_bucket)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in this graph, the distribution of travelers are roughly the same between their gender counterparts and age groups. 

#Countries data

```{r}
countries
```

All the possible predictable country destinations are listed here. As the origin is the US, the distances are the distance between the US and the country of destination. In addition, the levenshtein distance should be the calculated edit distance of language similiarities. It might be interesting to feature engineer something using the latitude and longitude to estimate temperature.