---
title: "New User Bookings Exploratory Analysis"
output: html_document
---

This is just some quick exploratory analysis done on the data provided. Anyways, the data is first loaded.

```{r}
#Packages
library(ggplot2)

#Data
dir = 'input/'
train = read.csv(paste0(dir,"train_users.csv"))
test = read.csv(paste0(dir,"test_users.csv"))
countries = read.csv(paste0(dir,"countries.csv"))
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
sessions  = read.csv(paste0(dir,"sessions.csv"))
```


#Training/Testing data

Create a quick function to show the head, tail, and number of na for each variable called

```{r}
sapply(train, function(y) sum(length(which(is.na(y)))))
```

The only column with NA values is age in train.

```{r}
sapply(test, function(y) sum(length(which(is.na(y)))))
```

The same can be seen in the test data set.

```{r}
table(train$age)
```

```{r}
table(test$age)
```

As can be seen in both the test and train set, the age variable will need some cleaning up. It is extremely unlikely that anyone 5 and under actually signed up and used the service. Because technically the rules state that you must be 18 to use the service, ages under 18 should be cleaned up. In addition, it appears that the ages variable has birth year data so those should be cleaned up accordingly. Although it would be hard to verify ages 100+, the oldest human recorded was around 117 so the outlier should be changed as well. Because AirBnb currently allows up to age 100, likely any data above are outliers however will remain due to the uncertainty of the true age.

```{r}
test$age[test$age > 150 & !is.na(test$age)] = 2015 - test$age[test$age > 150 & !is.na(test$age)]
train$age[train$age > 150 & !is.na(train$age)] = 2015 - train$age[train$age > 150 & !is.na(train$age)]
test$age[test$age > 115 & !is.na(test$age)] = 115
train$age[train$age > 115 & !is.na(train$age)] = 115
test$age[test$age < 18 & !is.na(test$age)] = 18
train$age[train$age < 18 & !is.na(train$age)] = 18

total = merge(train,test, all.x = TRUE, all.y=TRUE)

ggplot(total,aes(age)) + geom_line(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Distribution of ages")
```

The distribution for ages appears right-skewed with the peak around 29. There is another peak for users past 100 which may be outliers due to how the signup systems may have changed overtime or people misrepresenting their age.


#Sessions
```{r}
#See which values have na
sapply(sessions, function(y) sum(length(which(is.na(y)))))
```

The only columns with NA values is secs_elapsed.

```{r}
##Analyze the na sec
na_secs = sessions[which(is.na(sessions$secs_elapsed)),]

table(na_secs$action)[order(table(na_secs$action))]
table(na_secs$action_type)[order(table(na_secs$action_type))]
table(na_secs$action_detail)[order(table(na_secs$action_detail))]
```

It appears the the secs_elapsed are primarily NA values during creating and post requests.

```{r}
#Look at values with recorded secs_elapsed
no_na = sessions[which(!is.na(sessions$secs_elapsed)),]
table(no_na$action)[order(table(no_na$action))]
table(no_na$action_type)[order(table(no_na$action_type))]
table(no_na$action_detail)[order(table(no_na$action_detail))]
```

The rest of the data appear to primarily record get requests.

```{r}
#Look at the actions
table(sessions$action)[order(table(sessions$action))]
```

In general, it appears most of the actions are get queries. There is an odd value which appears to be blank done 51846 times.

```{r}
#Look at the action types
table(sessions$action_type)[order(table(sessions$action_type))]
```

The action type are similar to the action dataset. Details also includes a blank variable type.

```{r}
#Look at the action details
table(sessions$action_detail)[order(table(sessions$action_detail))]
```

The action follows the trend from the previous related variables and includes a blank variable.


```{r}
#Examine the blanks
table(sessions$action_detail[sessions$action_type == ""])[order(table(sessions$action_detail[sessions$action_type == ""]))]
table(sessions$action_type[sessions$action_detail == ""])[order(table(sessions$action_type[sessions$action_detail == ""]))]
```
The blanks can be seen in both action detail and type variables.

```{r}
table(sessions$action[sessions$action_detail == ""])[order(table(sessions$action[sessions$action_detail == ""]))]
table(sessions$device_type[sessions$action_detail == ""])[order(table(sessions$device_type[sessions$action_detail == ""]))]
ggplot(sessions[sessions$action == "",],aes(secs_elapsed)) + geom_line(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Seconds elapsed for blank action types")

```
It looks like this blank action happens frequently when searching and getting infomration from the show action. It appears that it happens on every device that tries to use it. In addition, time elapsed appears right skewed.

```{r}
##go take the top 10 actions
ggplot(sessions,aes(action)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")


```



```{r}
#See the counts of action types
ggplot(sessions,aes(action_type)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Counts of action type recorded by sessions")
```

The most popular action type is views, followed by data and click. There appears to be a fair amount of unknown action types being recorded.

```{r}
##Takes the top 10 action detail info
top_details = aggregate(sessions, by=list(sessions$action_detail), function(x) length(x))
top_details = top_details[,c(1,5)]
top_details = top_details[order(top_details$action_detail, na.last=FALSE),]
ggplot(tail(top_details,10),aes(x=Group.1, y=action_detail)) + geom_bar(stat="identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Top 10 Action Details")
```

The most popular action detailed type seems to be viewing search results. This makes sense as the service is heavily dependent on finding the right place to stay. In addition, the 2nd-4th most popular are p3, blank, and unknown. 

```{r}
#Show the counts of device types
ggplot(sessions,aes(device_type)) + geom_bar(stat="bin") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Counts of device types")
```
It appears that primarily, users using the service are still using desktops. There is a good amount of mobile users distributed primarily as iOS vs Android with microsoft and blackberry as uncommon. There also appears to be some mystery androidos derivatives using the service.


```{r}
#Look at the average time per action detail
detailed_avg = aggregate(sessions, by=list(sessions$action_detail), function(x) mean(as.numeric(x[6]), na.rm=TRUE))
detailed_avg = detailed_avg[,c(1,7)]
names(detailed_avg) = c("action_detail", "secs_elapsed")
detailed_avg[order(detailed_avg$secs_elapsed, na.last=FALSE),]
```
As can be seen in the averages of seconds elapsed by action, the times that have the highest average seconds elapsed are related to actions that would require large amount of time such as deciding on a password or deciding to book.




#Age data

The age data shows age bucket groups of users going to certain country destinations in 2015. 

```{r}
#switch 100+ age to 99+
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
levels(age$age_bucket) = c(levels(age$age_bucket), "99+")
age$age_bucket[age$age_bucket == "100+"] = "99+"

#plot destination vs population with gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Total population in thousands going to destinations separated by gender")
```

This shows that the gender distribution towards countries are roughly the same. The most population destination is within the US with the second most being Germany.

```{r}
#plot age vs population
ggplot(age,aes(x=age_bucket, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in the graph, there is an increase in travelers using the service in their 20s and begins to dip in the ages of 30-50. There is an upsurge in travel again in the travelers that are in their 50s and and begins to steadily decrease with age. There appears to be more males using the service that are in their mid-30s and below however begins to even out and become eclipsed by their female counterpart afterwards.

```{r}
#plot destinatoin vs population with alpha of age and gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender, alpha=age_bucket)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in this graph, the distribution of travelers are roughly the same between their gender counterparts and age groups. 




#Countries data

```{r}
countries
```

All the possible predictable country destinations are listed here. As the origin is the US, the distances are the distance between the US and the country of destination. In addition, the levenshtein distance should be the calculated edit distance of language similiarities. It might be interesting to feature engineer something using the latitude and longitude to estimate temperature.



#Merging the data with sessions
The training and testing data is merged together to look at the data that will be used for prediction

```{r}
#needs to do merge all
#Not sure if x or y but ultimately will have to predit based on unique user id at this rate
train = merge(train, sessions, by.x="id", by.y="user_id", all.y = TRUE)
test = merge(test, sessions, by.x="id", by.y="user_id")
```

