---
title: "New User Bookings Exploratory Analysis"
output: html_document
---

This is just some quick exploratory analysis done on the data provided. Anyways, the data is first loaded.

```{r}
dir = 'input/'
train = read.csv(paste0(dir,"train_users.csv"))
test = read.csv(paste0(dir,"test_users.csv"))
countries = read.csv(paste0(dir,"countries.csv"))
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
sessions  = read.csv(paste0(dir,"sessions.csv"))
```

The training and testing data is merged together for future analysis

```{r}
train = merge(train, sessions, by.x="id", by.y="user_id")
test = merge(test, sessions, by.x="id", by.y="user_id")
```

#Training/Testing data

Create a quick function to show the head, tail, and number of na for each variable called

```{r}
printData = function(df, var){
    print(table(is.na(df[var])))
    print(head(table(df[var])))
    print(tail(table(df[var])))
}
```


#Age data

```{r}

```

#Countries data

```{r}
countries
```

All the possible predictable country destinations are listed here. As the origin is the US, the distances are the distance between the US and the country of destination.