---
title: "New User Bookings Exploratory Analysis"
output: html_document
---

This is just some quick exploratory analysis done on the data provided. Anyways, the data is first loaded.

```{r}
#Packages
library(ggplot2)

#Data
dir = 'input/'
train = read.csv(paste0(dir,"train_users.csv"))
test = read.csv(paste0(dir,"test_users.csv"))
countries = read.csv(paste0(dir,"countries.csv"))
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
sessions  = read.csv(paste0(dir,"sessions.csv"))
```

The training and testing data is merged together to look at the data that will be used for prediction

```{r}
train = merge(train, sessions, by.x="id", by.y="user_id")
test = merge(test, sessions, by.x="id", by.y="user_id")
```

#Training/Testing data

Create a quick function to show the head, tail, and number of na for each variable called

```{r}
printData = function(df, var){
    print(table(is.na(df[var])))
    print(head(table(df[var])))
    print(tail(table(df[var])))
}
```


#Age data

The age data shows age bucket groups of users going to certain country destinations in 2015. 

```{r}
#switch 100+ age to 99+
age = read.csv(paste0(dir, "age_gender_bkts.csv"))
levels(age$age_bucket) = c(levels(age$age_bucket), "99+")
age$age_bucket[age$age_bucket == "100+"] = "99+"

#plot destination vs population with gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Total population in thousands going to destinations separated by gender")
```

This shows that the gender distribution towards countries are roughly the same. The most population destination is within the US with the second most being Germany.

```{r}
#plot age vs population
ggplot(age,aes(x=age_bucket, y=population_in_thousands, fill=gender)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in the graph, there is an increase in travelers using the service in their 20s and begins to dip in the ages of 30-50. There is an upsurge in travel again in the travelers that are in their 50s and and begins to steadily decrease with age. There appears to be more males using the service that are in their mid-30s and below however begins to even out and become eclipsed by their female counterpart afterwards.

```{r}
#plot destinatoin vs population with alpha of age and gender as fill
ggplot(age,aes(x=country_destination, y=population_in_thousands, fill=gender, alpha=age_bucket)) + geom_bar(stat="identity", position="dodge") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(title="Population of travelers in thousands traveling in 2015")
```

As can be seen in this graph, the distribution of travelers are roughly the same between their gender counterparts and age groups. 

#Countries data

```{r}
countries
```

All the possible predictable country destinations are listed here. As the origin is the US, the distances are the distance between the US and the country of destination. In addition, the levenshtein distance should be the calculated edit distance of language similiarities. It might be interesting to feature engineer something using the latitude and longitude to estimate temperature.